---
title: "Exploratory Data Analysis"
author: "QingZhang, Suraj Kumar, Xin Zhang, Xinyue Shu, Yao Xiao"
date: "2021/7/18"
output: github_document

number_sections: yes
always_allow_html: true

fig_caption: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(moderndive)
library(skimr)
library(kableExtra)
library(gridExtra)
library(GGally)
library(infer)
library(broom)
library(ggfortify)
library(jtools)
library(AER)
library(car)
library(janitor)
```


```{r dataset, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
dataset4<-read.csv("D:/02 - Data Analytics/08 - Data Analysis Skills/GroupProject2/dataset4.csv")
#loaded the dataset
dataset4<-dataset4%>%
  select(-Region) #delete unnecessary variable
dataset4[sapply(dataset4,is.character)]<-lapply(dataset4[sapply(dataset4,is.character)],as.factor)
#modified character variable to factor
dataset4["Electricity"]<-lapply(dataset4["Electricity"],as.factor)
#modified Electricity variable to factor
```

# Introduction {#sec:Intro}

The society is composed of one family, and the study of the population of family members is of great significance to the population of the study area. Understanding the deciding factors that affect the number of family member numbers is helpful to understand the population situation in the area, predict future population development trends and make timely policy adjustments. The Philippine government conducts surveys on household income and expenditure every three years to understand the living conditions of residents, dataset comes from the FIES (Family Income and Expenditure Survey) recorded in the Philippines. 

In some past studies, we found that some factors may affect the number of family members. This research studied 2122 families in Soccsksargen district and collected the data of total household income, total food expenditure, household head sex, household head age, type of household, total number of family members, house floor area, house age, number of bedrooms and electricity, the purpose is to find the relationship between the number of family members and other variables. 
This report focuses on a different analysis level through summaries, boxplots, and general linear model. Section \ref{sec:EDA} consists of an exploratory data analysis of number of family members and explores the potential relationship between member numbers and other variables. Section \ref{sec:FDA} contains the results from fitting a generalized linear model to the data, as well as the assessment of the model assumptions. Concluding remarks are given in Section \ref{sec:Conc}.

# Exploratory data analysis {#sec:EDA}
The data set contains 2122 observations with no missing value, and records a total of 10 variables, including three categorical variables and seven numerical variables. \ref{tab:Summaries1} shows the general information of the data set, including mean, standard deviation, quartile, etc. For making it clearer, this part of the analysis will covertly analyze different categorical variables one by one.

```{r, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE,fig.cap="\\label{tab:summary} Summary Statistics of Data Set", fig.pos='H'}
skim(dataset4)
#produce summary statistics of all variables
```

